---
title: "The Econometrician’s Lens — Issue #1"
subtitle: "Pakistan’s Trade Gap: Data, Politics, and Perception"
author: "Zahid Asghar"
date: "2025-10-18"
categories: [Pakistan, Trade, Econometrics, Policy]
format:
  html:
    theme: cosmo
    toc: true
    toc-location: left
    number-sections: true
    smooth-scroll: true
    anchor-sections: true
    code-copy: true
    code-link: true
execute:
  echo: true
  warning: false
  message: false
editor: visual
---

::: callout-note
**Welcome to the launch edition.**\
*A weekly visual, data‑driven look at Pakistan’s economy — where evidence meets narrative.*
:::

# Opening Hook

> When a trade gap becomes a political storm, data becomes the first casualty.

Pakistan’s recent trade‑data debate (the “\$11B vs. \$6B” import gap) exposed structural weaknesses in transparency, reconciliation, and statistical capacity. This launch issue explains the **data story**, suggests how an **econometrician would test claims**, and proposes **practical fixes** that improve evidence‑based policymaking.

------------------------------------------------------------------------

# Data Pulse — Visual Insight

**Title:** *Imports vs. Exports — The Quiet Divergence (FY2022–FY2025)*

Use the template below. Replace `pak_trade_data.csv` with your dataset (PBS/SBP monthly).\
Expected columns: `Month, Imports, Exports` where `Month` is ISO date (YYYY-MM-01) and values are in **million USD**.

```{r}
#| label: trade-plot
#| echo: true
#| fig-cap: "Pakistan Trade Flows (FY2022–FY2025) — template"
# Install once if needed:
# install.packages(c("readr","ggplot2","scales"))
library(readr)
library(ggplot2)
library(scales)

df <- read_csv("pak_trade_data.csv")  # <- replace with your path
df$Month <- as.Date(df$Month)

ggplot(df, aes(x = Month)) +
  geom_line(aes(y = Imports, linetype = "Imports"), linewidth = 1.1) +
  geom_line(aes(y = Exports, linetype = "Exports"), linewidth = 1.1) +
  scale_y_continuous(labels = label_number_si()) +
  labs(title = "Pakistan Trade Flows (FY2022–FY2025)",
       y = "Million USD", x = "Month", linetype = NULL) +
  theme_minimal(base_size = 13)
```

**Insight (1–2 lines):** Despite some reporting improvements in 2024–25, the **trade deficit remains structurally persistent**; remittances and valuation effects can mask deeper competitiveness issues.

------------------------------------------------------------------------

# Deep Dive — What the IMF Found (and Why It Matters)

-   The controversy centers on **possible over/under‑statement of imports**, **valuation gaps**, and **lags** across reporting systems.\
-   Divergences can emerge from: customs declarations, exchange‑rate timing, informal channels, and methodological differences between **PBS** (goods trade) and **SBP** (balance‑of‑payments).\
-   **Governance lesson:** Transparency in data systems is not cosmetic — it’s a **policy instrument**. Open, reconcilable data reduces uncertainty for firms, markets, and citizens.

> *Credibility is a macroeconomic variable.*

------------------------------------------------------------------------

# Econometrics in Action — Testing a Policy Claim

**Question:** Did import restrictions reduce domestic prices?

Two quick templates you can adapt.

### Stata (DiD)

``` stata
* price_it = β0 + β1 post_t + β2 treated_i + β3 (post_t×treated_i) + γX_it + ε_it
reghdfe price i.post##i.treated controls, absorb(id time) vce(robust)
esttab, se ar2
```

### Python (statsmodels)

``` python
import statsmodels.formula.api as smf
model = smf.ols("price ~ post * treated + controls", data=df).fit(cov_type="HC3")
print(model.summary())
```

**Interpretation tip:** β₃ (the interaction) captures the differential post‑policy change for the treated group versus control, controlling for confounders and fixed effects where applicable.

------------------------------------------------------------------------

# Policy & Governance Note — Reconciliation Wins

-   **Digitize and reconcile** customs→PBS→SBP flows with a public changelog.\
-   Publish **method notes** for adjustments/deflators, and maintain a **versioned dataset**.\
-   Build a **monthly discrepancy dashboard** (PBS vs SBP) with flags when gaps exceed thresholds.

**Schematic (suggested for future graphic):**\
Customs → PBS (goods) → SBP (BoP) → IMF Reporting → Public Portal

------------------------------------------------------------------------

# Tool of the Week — Shiny Reconciliation Starter

Prototype a quick validator to compare PBS and SBP series.

```{r}
#| label: shiny-starter
# Minimal Shiny pseudo‑starter (expand in your app):
# install.packages("shiny")
library(shiny)

ui <- fluidPage(
  titlePanel("Trade Data Reconciliation — Demo"),
  sidebarLayout(
    sidebarPanel(
      fileInput("pbs", "Upload PBS CSV"),
      fileInput("sbp", "Upload SBP CSV")
    ),
    mainPanel(
      plotOutput("gapPlot")
    )
  )
)

server <- function(input, output, session) {
  output$gapPlot <- renderPlot({
    # Read, align dates, compute gap, then plot.
    # Placeholder: add your logic here.
    plot.new(); title("Gap plot will render here")
  })
}

# shinyApp(ui, server)  # Uncomment to run locally
```

------------------------------------------------------------------------

# What to Watch Next Week

-   Data releases: CPI, remittances, FX reserves, and any policy circulars impacting imports.\
-   Narrative risk: headlines vs. methodology — expect revisions and reconciliation notes.

------------------------------------------------------------------------

# One‑Line Insight

> When numbers become narratives, **transparency becomes governance**.

------------------------------------------------------------------------

# About & Call to Action

If you find this useful, share it with colleagues in policy, academia, and industry.\
I’m building open, replicable workflows in **R/Python/Stata** and consulting on **econometrics, impact evaluation, and data systems**.

-   **Subscribe:** Get the next issue in your inbox.\
-   **Collaborate:** Commission a brief, dashboard, or evaluation design.\
-   **Contact:** zahid \[at\] email — LinkedIn/X (DMs open).
