---
title: "Interactive AI Research Assistant"
subtitle: "Integrating AI into Academic Research Workflow"
author: "Economics Research Lab"
date: today
format: 
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    code-tools: true
    embed-resources: true
    grid:
      sidebar-width: 300px
      body-width: 900px
      margin-width: 300px
execute:
  echo: true
  warning: false
  message: false
---

# Interactive AI Research Assistant in Quarto

This document demonstrates how to integrate AI assistance directly into your research workflow using Quarto's interactive capabilities.

## Setup and Libraries

```{r}
#| label: setup
#| include: false

# Load required libraries
library(tidyverse)
library(DT)
library(plotly)
library(reactable)
library(htmltools)
library(modelsummary)

# Set global options
options(digits = 3)
knitr::opts_chunk$set(
  fig.width = 10,
  fig.height = 6,
  dpi = 300,
  warning = FALSE,
  message = FALSE
)
```

## Research Context Manager

```{r}
#| label: research-context
#| echo: true

# Create a research context structure
research_context <- list(
  project_title = "Impact of AI on Labor Market Dynamics",
  methodology = "Panel Data Analysis",
  data_source = "Administrative Records + Survey Data",
  research_questions = c(
    "How does AI adoption affect employment levels?",
    "What are the skill-biased effects of AI implementation?",
    "How do labor market outcomes vary by industry?"
  ),
  analysis_tools = c("R", "Python", "Stata"),
  status = "Data Collection Phase"
)

# Display research context
research_context
```

## Interactive AI Assistant Interface

::: {.panel-tabset}

## AI Assistant

```{=html}
<div id="ai-assistant" style="
  border: 2px solid #e3f2fd;
  border-radius: 10px;
  padding: 20px;
  margin: 20px 0;
  background: linear-gradient(135deg, #f8f9fa 0%, #e3f2fd 100%);
  box-shadow: 0 4px 6px rgba(0,0,0,0.1);
">
  <h3 style="color: #1565c0; margin-bottom: 15px;">
    🤖 AI Research Assistant
  </h3>
  
  <div style="margin-bottom: 15px;">
    <strong>Quick Assistance:</strong>
    <div style="display: flex; gap: 10px; margin: 10px 0; flex-wrap: wrap;">
      <button onclick="setAssistanceType('methodology')" style="
        background: #2196f3;
        color: white;
        border: none;
        padding: 8px 16px;
        border-radius: 20px;
        cursor: pointer;
        font-size: 12px;
      ">📊 Methodology</button>
      <button onclick="setAssistanceType('analysis')" style="
        background: #4caf50;
        color: white;
        border: none;
        padding: 8px 16px;
        border-radius: 20px;
        cursor: pointer;
        font-size: 12px;
      ">📈 Analysis</button>
      <button onclick="setAssistanceType('code')" style="
        background: #ff9800;
        color: white;
        border: none;
        padding: 8px 16px;
        border-radius: 20px;
        cursor: pointer;
        font-size: 12px;
      ">💻 Code</button>
      <button onclick="setAssistanceType('writing')" style="
        background: #9c27b0;
        color: white;
        border: none;
        padding: 8px 16px;
        border-radius: 20px;
        cursor: pointer;
        font-size: 12px;
      ">✍️ Writing</button>
    </div>
  </div>
  
  <div style="margin-bottom: 15px;">
    <textarea id="research-query" placeholder="Ask about methodology, analysis, coding, or writing assistance..." style="
      width: 100%;
      height: 80px;
      padding: 10px;
      border: 2px solid #e0e0e0;
      border-radius: 5px;
      font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
      resize: vertical;
    "></textarea>
  </div>
  
  <div style="display: flex; gap: 10px; margin-bottom: 15px;">
    <button onclick="getAIAssistance()" style="
      background: #1565c0;
      color: white;
      border: none;
      padding: 10px 20px;
      border-radius: 5px;
      cursor: pointer;
      font-weight: bold;
    ">Get AI Assistance</button>
    <button onclick="clearResponse()" style="
      background: #757575;
      color: white;
      border: none;
      padding: 10px 20px;
      border-radius: 5px;
      cursor: pointer;
    ">Clear</button>
  </div>
  
  <div id="ai-response" style="
    background: white;
    border: 1px solid #e0e0e0;
    border-radius: 5px;
    padding: 15px;
    min-height: 100px;
    max-height: 400px;
    overflow-y: auto;
    font-family: 'Segoe UI', Tahoma, Geneva, Verdana, sans-serif;
    line-height: 1.6;
    display: none;
  ">
    <div id="response-content"></div>
  </div>
</div>
```

## Usage Guide

The AI assistant above provides specialized help for economics research:

- **📊 Methodology**: Research design, sampling, causality
- **📈 Analysis**: Statistical techniques, econometric models
- **💻 Code**: R/Python implementation, debugging
- **✍️ Writing**: Structure, argumentation, academic style

### Example Queries:

- "Help me design an IV strategy for studying AI adoption effects"
- "I need R code for difference-in-differences with multiple time periods"
- "How should I structure the methodology section for a labor economics paper?"

:::

<script>
function setAssistanceType(type) {
  const textarea = document.getElementById('research-query');
  const templates = {
    methodology: "Help me design a methodology for studying [research topic]. I need guidance on [specific aspect].",
    analysis: "I have [describe your data] and want to perform [type of analysis]. What R/Python approach should I take?",
    code: "I need help with R code for [specific task]. My data structure includes [describe variables].",
    writing: "I'm working on [paper section] and need help structuring my argument about [topic]."
  };
  textarea.value = templates[type];
  textarea.focus();
}

function getAIAssistance() {
  const query = document.getElementById('research-query').value;
  const responseDiv = document.getElementById('ai-response');
  const contentDiv = document.getElementById('response-content');
  
  if (!query.trim()) {
    alert('Please enter a research question or request.');
    return;
  }
  
  // Show loading state
  responseDiv.style.display = 'block';
  contentDiv.innerHTML = `
    <div style="text-align: center; padding: 20px;">
      <div style="display: inline-block; width: 20px; height: 20px; border: 3px solid #f3f3f3; border-top: 3px solid #3498db; border-radius: 50%; animation: spin 1s linear infinite;"></div>
      <p style="margin-top: 10px; color: #666;">Generating AI assistance...</p>
    </div>
  `;
  
  // Simulate AI response (in real implementation, this would call Claude API)
  setTimeout(() => {
    const responses = {
      methodology: generateMethodologyResponse(query),
      analysis: generateAnalysisResponse(query),
      code: generateCodeResponse(query),
      writing: generateWritingResponse(query),
      default: generateDefaultResponse(query)
    };
    
    const responseType = detectResponseType(query);
    contentDiv.innerHTML = responses[responseType] || responses.default;
  }, 2000);
}

function detectResponseType(query) {
  const q = query.toLowerCase();
  if (q.includes('methodology') || q.includes('design') || q.includes('approach')) return 'methodology';
  if (q.includes('analysis') || q.includes('statistical') || q.includes('model')) return 'analysis';
  if (q.includes('code') || q.includes('r ') || q.includes('python')) return 'code';
  if (q.includes('writing') || q.includes('paper') || q.includes('draft')) return 'writing';
  return 'default';
}

function generateMethodologyResponse(query) {
  return `
    <div style="background: #e8f5e8; padding: 15px; border-radius: 5px; margin-bottom: 15px;">
      <h4 style="color: #2e7d32; margin-bottom: 10px;">🎯 Methodology Guidance</h4>
      <p><strong>Research Design Recommendation:</strong></p>
      <p>Based on your query about "${query.substring(0, 100)}...", I recommend considering a mixed-methods approach that combines quantitative panel data analysis with qualitative interviews.</p>
      
      <p><strong>Key Considerations:</strong></p>
      <ul>
        <li>Sample size and power analysis requirements</li>
        <li>Potential confounding variables and control strategies</li>
        <li>Temporal dynamics and causality assessment</li>
        <li>External validity and generalizability</li>
      </ul>
      
      <p><strong>Next Steps:</strong></p>
      <ol>
        <li>Develop detailed sampling strategy</li>
        <li>Create data collection instruments</li>
        <li>Plan analytical framework</li>
        <li>Consider ethical clearance requirements</li>
      </ol>
    </div>
  `;
}

function generateAnalysisResponse(query) {
  return `
    <div style="background: #e3f2fd; padding: 15px; border-radius: 5px; margin-bottom: 15px;">
      <h4 style="color: #1565c0; margin-bottom: 10px;">📊 Analysis Guidance</h4>
      <p><strong>Analytical Approach:</strong></p>
      <p>For your research question "${query.substring(0, 100)}...", I recommend starting with exploratory data analysis followed by fixed-effects panel regression.</p>
      
      <p><strong>Suggested Analysis Pipeline:</strong></p>
      <ol>
        <li><strong>Descriptive Statistics:</strong> Summary statistics, correlation matrices</li>
        <li><strong>Visualization:</strong> Time series plots, cross-sectional comparisons</li>
        <li><strong>Model Selection:</strong> Fixed vs random effects testing</li>
        <li><strong>Robustness Checks:</strong> Alternative specifications, subgroup analysis</li>
      </ol>
      
      <p><strong>R Packages to Consider:</strong></p>
      <ul>
        <li><code>plm</code> - Panel data econometrics</li>
        <li><code>lfe</code> - Linear group fixed effects</li>
        <li><code>modelsummary</code> - Publication-ready tables</li>
        <li><code>ggplot2</code> - Data visualization</li>
      </ul>
    </div>
  `;
}

function generateCodeResponse(query) {
  return `
    <div style="background: #fff3e0; padding: 15px; border-radius: 5px; margin-bottom: 15px;">
      <h4 style="color: #f57c00; margin-bottom: 10px;">💻 Code Assistance</h4>
      <p><strong>Code Solution:</strong></p>
      <p>Here's R code to address your request about "${query.substring(0, 100)}...":</p>
      
      <pre style="background: #2d3748; color: #e2e8f0; padding: 15px; border-radius: 5px; overflow-x: auto;"><code># Load required libraries
library(tidyverse)
library(plm)
library(modelsummary)

# Load and prepare data
data <- read_csv("your_data.csv") %>%
  mutate(
    time_period = as.factor(year),
    entity_id = as.factor(id)
  )

# Run panel regression
model <- plm(outcome ~ treatment + controls, 
             data = data, 
             index = c("entity_id", "time_period"),
             model = "within")

# Generate summary table
modelsummary(model, 
             stars = TRUE,
             gof_map = c("nobs", "r.squared"))
</code></pre>
      
      <p><strong>Key Points:</strong></p>
      <ul>
        <li>Ensure proper data structure with entity and time identifiers</li>
        <li>Consider clustering standard errors at appropriate level</li>
        <li>Validate model assumptions through diagnostic plots</li>
      </ul>
    </div>
  `;
}

function generateWritingResponse(query) {
  return `
    <div style="background: #f3e5f5; padding: 15px; border-radius: 5px; margin-bottom: 15px;">
      <h4 style="color: #7b1fa2; margin-bottom: 10px;">✍️ Writing Assistance</h4>
      <p><strong>Writing Guidance:</strong></p>
      <p>For your writing query "${query.substring(0, 100)}...", here's a structured approach:</p>
      
      <p><strong>Section Structure:</strong></p>
      <ol>
        <li><strong>Opening:</strong> State main argument clearly</li>
        <li><strong>Literature Context:</strong> Position within existing research</li>
        <li><strong>Evidence:</strong> Present empirical support</li>
        <li><strong>Implications:</strong> Discuss broader significance</li>
      </ol>
      
      <p><strong>Academic Writing Tips:</strong></p>
      <ul>
        <li>Use active voice where possible</li>
        <li>Ensure logical flow between paragraphs</li>
        <li>Support claims with evidence</li>
        <li>Maintain consistent terminology</li>
      </ul>
      
      <p><strong>Sample Paragraph Structure:</strong></p>
      <p style="font-style: italic; background: #fafafa; padding: 10px; border-left: 4px solid #7b1fa2;">
        "The evidence suggests that [main finding]. This result is consistent with [theoretical framework] and extends previous work by [specific contribution]. The implications for [broader context] are significant because [explanation]."
      </p>
    </div>
  `;
}

function generateDefaultResponse(query) {
  return `
    <div style="background: #f5f5f5; padding: 15px; border-radius: 5px; margin-bottom: 15px;">
      <h4 style="color: #424242; margin-bottom: 10px;">🔍 Research Assistance</h4>
      <p><strong>General Guidance:</strong></p>
      <p>Thank you for your query: "${query.substring(0, 100)}..."</p>
      
      <p>I can help you with:</p>
      <ul>
        <li><strong>Research Design:</strong> Methodology selection and study design</li>
        <li><strong>Data Analysis:</strong> Statistical techniques and interpretation</li>
        <li><strong>Code Development:</strong> R/Python implementation</li>
        <li><strong>Academic Writing:</strong> Structure and argumentation</li>
      </ul>
      
      <p>Please feel free to ask more specific questions about any of these areas, and I'll provide detailed guidance tailored to your research needs.</p>
    </div>
  `;
}

function clearResponse() {
  document.getElementById('ai-response').style.display = 'none';
  document.getElementById('research-query').value = '';
}

// Add CSS for spinner animation
const style = document.createElement('style');
style.textContent = `
  @keyframes spin {
    0% { transform: rotate(0deg); }
    100% { transform: rotate(360deg); }
  }
`;
document.head.appendChild(style);
</script>
```

## Interactive Data Analysis Example

```{r}
#| label: sample-analysis
#| echo: true

# Generate sample data for demonstration
set.seed(42)
n <- 1000

sample_data <- tibble(
  id = 1:n,
  industry = sample(c("Manufacturing", "Services", "Technology", "Finance"), n, replace = TRUE),
  ai_adoption = rbinom(n, 1, 0.3),
  firm_size = rlnorm(n, 4, 1),
  employment_change = rnorm(n, 
    mean = ifelse(ai_adoption == 1, -0.05, 0.02), 
    sd = 0.15),
  productivity_change = rnorm(n, 
    mean = ifelse(ai_adoption == 1, 0.12, 0.03), 
    sd = 0.20)
)

# Display interactive table
sample_data %>%
  slice_head(n = 100) %>%
  reactable(
    columns = list(
      employment_change = colDef(format = colFormat(digits = 3)),
      productivity_change = colDef(format = colFormat(digits = 3)),
      firm_size = colDef(format = colFormat(digits = 0))
    ),
    searchable = TRUE,
    filterable = TRUE,
    striped = TRUE,
    highlight = TRUE,
    bordered = TRUE,
    theme = reactableTheme(
      borderColor = "#dfe2e5",
      stripedColor = "#f6f8fa",
      highlightColor = "#f0f5ff",
      cellPadding = "8px 12px"
    )
  )
```

## Interactive Visualization

```{r}
#| label: interactive-plot
#| echo: true

# Create interactive visualization
p <- sample_data %>%
  ggplot(aes(x = firm_size, y = employment_change, 
             color = factor(ai_adoption), 
             text = paste("Industry:", industry,
                         "<br>Firm Size:", round(firm_size, 0),
                         "<br>Employment Change:", round(employment_change, 3)))) +
  geom_point(alpha = 0.6, size = 2) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  scale_color_manual(values = c("0" = "#ff7f00", "1" = "#1f77b4"),
                     labels = c("No AI", "AI Adopted")) +
  labs(
    title = "AI Adoption and Employment Effects by Firm Size",
    x = "Firm Size (log scale)",
    y = "Employment Change (%)",
    color = "AI Adoption"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    legend.position = "bottom"
  )

# Convert to interactive plot
ggplotly(p, tooltip = "text") %>%
  layout(
    title = list(text = "AI Adoption and Employment Effects by Firm Size", 
                 font = list(size = 16)),
    hovermode = "closest"
  )
```

## Analysis Summary with AI Integration

```{r}
#| label: analysis-summary
#| echo: true

# Perform regression analysis
model_employment <- lm(employment_change ~ ai_adoption + log(firm_size) + industry, 
                      data = sample_data)

model_productivity <- lm(productivity_change ~ ai_adoption + log(firm_size) + industry, 
                        data = sample_data)

# Create summary table
modelsummary(
  list("Employment Change" = model_employment,
       "Productivity Change" = model_productivity),
  stars = TRUE,
  gof_map = c("nobs", "r.squared", "adj.r.squared"),
  title = "Impact of AI Adoption on Labor Market Outcomes"
)
```

## Research Workflow Integration

### 1. Literature Review Assistant

::: {.callout-note}
## 📚 Literature Review Integration

Use the AI assistant above to:

- Generate literature search strategies
- Identify key themes and gaps
- Structure literature review sections
- Suggest theoretical frameworks

**Example Query:** "Help me structure a literature review on AI and labor markets, focusing on skill-biased technological change"
:::

### 2. Methodology Development

::: {.callout-tip}
## 🔬 Methodology Assistant

Get guidance on:

- Research design selection
- Sampling strategies
- Data collection methods
- Analytical approaches

**Example Query:** "I want to study firm-level AI adoption effects using panel data. What methodology should I consider?"
:::

### 3. Code Development Support

::: {.callout-important}
## 💻 Code Assistant

Receive help with:

- R/Python implementation
- Statistical modeling
- Data visualization
- Debugging and optimization

**Example Query:** "I need R code for instrumental variable estimation with firm-level clustered standard errors"
:::

### 4. Writing and Publication Support

::: {.callout-warning}
## ✍️ Writing Assistant

Get support for:

- Paper structure and argumentation
- Results interpretation
- Abstract and conclusion writing
- Response to reviewer comments

**Example Query:** "Help me write the discussion section for my paper on AI adoption and employment effects"
:::

## Implementation Notes

### Setting Up Claude API Integration

To make this fully functional with actual AI responses, you'll need to:

1. **Get Claude API access** from Anthropic
2. **Replace the mock responses** with actual API calls
3. **Add proper error handling** and security measures
4. **Implement conversation memory** for context continuity

### Example API Integration Code

```javascript
// Add to your Quarto document's JavaScript section
async function callClaudeAPI(query, context) {
  const response = await fetch('https://api.anthropic.com/v1/messages', {
    method: 'POST',
    headers: {
      'Content-Type': 'application/json',
      'Authorization': 'Bearer YOUR_API_KEY',
      'anthropic-version': '2023-06-01'
    },
    body: JSON.stringify({
      model: 'claude-3-sonnet-20240229',
      max_tokens: 1000,
      messages: [{
        role: 'user',
        content: `Research Context: ${JSON.stringify(context)}
        
        Query: ${query}
        
        Please provide specific, actionable advice for this economics research query.`
      }]
    })
  });
  
  const data = await response.json();
  return data.content[0].text;
}
```

## Benefits of This Approach

1. **Seamless Integration**: AI assistance embedded directly in your research documents
2. **Context Preservation**: Research context maintained throughout the document
3. **Reproducible Research**: All queries and responses documented
4. **Collaborative**: Easy to share with colleagues and co-authors
5. **Version Control**: Works with Git and other version control systems

## Next Steps

1. **Customize the assistant** for your specific research areas
2. **Add more sophisticated prompts** for different types of queries
3. **Integrate with your existing R/Python workflows**
4. **Deploy as interactive web documents** for broader access

---

*This interactive research assistant demonstrates the power of combining AI capabilities with Quarto's rich document features for enhanced academic research workflows.*