---
title: "Pakistan's Tax Assault on the Salaried Class: A Decade of Exploitation"
subtitle: "The Devastating Impact of Progressive Taxation Amid Hyperinflation"
author: "Zahid Asghar"
date: "`r Sys.Date()`"
format: 
  html:
    theme: cosmo
    toc: true
    toc-depth: 3
    code-fold: true
    warning: false
    message: false
    fig-width: 10
    fig-height: 6
execute:
  echo: false
  warning: false
  message: false
---

```{r setup}
#| echo: false
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(ggplot2)
library(plotly)
library(DT)

# Set global options
options(scipen = 999)
```



This analysis reveals the **shocking reality** of Pakistan's tax policy on salaried individuals over the past five years. Combined with one of the highest historical inflation rates reaching **37%** in 2023, the salaried class faces a double burden: higher nominal tax rates and severely eroded purchasing power.

## Tax Calculation Functions

```{r tax-functions}
#| echo: true

# Tax calculation functions for different years
calculate_tax_2025 <- function(income) {
  case_when(
    income <= 600000 ~ 0,
    income <= 1200000 ~ (income - 600000) * 0.05,
    income <= 2200000 ~ 30000 + (income - 1200000) * 0.15,
    income <= 3200000 ~ 180000 + (income - 2200000) * 0.25,
    income <= 4100000 ~ 430000 + (income - 3200000) * 0.30,
    TRUE ~ 700000 + (income - 4100000) * 0.35
  )
}

calculate_tax_2024 <- function(income) {
  case_when(
    income <= 600000 ~ 0,
    income <= 1200000 ~ (income - 600000) * 0.025,
    income <= 2400000 ~ 15000 + (income - 1200000) * 0.125,
    income <= 3600000 ~ 165000 + (income - 2400000) * 0.225,
    income <= 6000000 ~ 435000 + (income - 3600000) * 0.275,
    TRUE ~ 1095000 + (income - 6000000) * 0.35
  )
}

calculate_tax_2020 <- function(income) {
  case_when(
    income <= 600000 ~ 0,
    income <= 1200000 ~ (income - 600000) * 0.05,
    income <= 1800000 ~ 30000 + (income - 1200000) * 0.10,
    income <= 2500000 ~ 90000 + (income - 1800000) * 0.15,
    income <= 3500000 ~ 195000 + (income - 2500000) * 0.175,
    income <= 5000000 ~ 370000 + (income - 3500000) * 0.20,
    income <= 8000000 ~ 670000 + (income - 5000000) * 0.225,
    income <= 12000000 ~ 1345000 + (income - 8000000) * 0.25,
    income <= 30000000 ~ 2345000 + (income - 12000000) * 0.275,
    income <= 50000000 ~ 7295000 + (income - 30000000) * 0.30,
    income <= 75000000 ~ 13295000 + (income - 50000000) * 0.325,
    TRUE ~ 21420000 + (income - 75000000) * 0.35
  )
}
```

## Inflation Impact Analysis

Pakistan like many other countries  has experienced devastating post-covid inflation over the past 5 years:

```{r inflation-data}
#| echo: true

# Pakistan's inflation rates (Consumer Price Index)
inflation_data <- data.frame(
  Year = c(2020, 2021, 2022, 2023, 2024, 2025),
  Inflation_Rate = c(10.7, 12.3, 19.9, 37.2, 23.1, 15.8),  # Approximate rates
  Cumulative_Inflation = c(1.0, 1.123, 1.346, 1.847, 2.273, 2.632)
)

# Real purchasing power calculator
real_value_2020 <- function(nominal_2025, cumulative_inflation = 2.632) {
  nominal_2025 / cumulative_inflation
}
```

::: {.callout-important}
## Critical Reality Check
A salary of **Rs. 1.5 million in 2025** has the same purchasing power as **Rs. 570,000 in 2020** due to cumulative inflation of **163%**!
:::

## The Brutal Reality: Tax Burden Analysis

```{r main-analysis}
# Define income levels representing different economic classes
income_levels <- c(800000, 1500000, 2500000, 3500000, 5000000)
income_labels <- c("Lower Middle", "Middle Class", "Upper Middle", "Professional", "High Earners")

# Create comprehensive analysis
tax_analysis <- data.frame(
  Income_Level = income_labels,
  Nominal_Income = income_levels,
  Real_Income_2020_Terms = round(real_value_2020(income_levels), 0),
  Tax_2020 = calculate_tax_2020(income_levels),
  Tax_2024 = calculate_tax_2024(income_levels),
  Tax_2025 = calculate_tax_2025(income_levels)
) %>%
  mutate(
    # Tax rates
    Rate_2020 = round((Tax_2020 / Nominal_Income) * 100, 2),
    Rate_2024 = round((Tax_2024 / Nominal_Income) * 100, 2),
    Rate_2025 = round((Tax_2025 / Nominal_Income) * 100, 2),
    
    # Percentage increases
    Increase_2024_vs_2020 = round(((Tax_2024 - Tax_2020) / Tax_2020) * 100, 1),
    Increase_2025_vs_2024 = round(((Tax_2025 - Tax_2024) / Tax_2024) * 100, 1),
    Increase_2025_vs_2020 = round(((Tax_2025 - Tax_2020) / Tax_2020) * 100, 1),
    
    # Real tax burden (inflation-adjusted)
    Real_Tax_Burden_Change = round(((Tax_2025 / 2.632) - Tax_2020) / Tax_2020 * 100, 1),
    
    # Disposable income after tax
    Disposable_2020 = Nominal_Income - Tax_2020,
    Disposable_2025 = Nominal_Income - Tax_2025,
    Real_Disposable_2025 = round((Nominal_Income - Tax_2025) / 2.632, 0),
    
    # Real disposable income change
    Real_Disposable_Change = round(((Real_Disposable_2025 - Disposable_2020) / Disposable_2020) * 100, 1)
  )

# Replace Inf and NaN values
tax_analysis[is.infinite(tax_analysis$Increase_2024_vs_2020), "Increase_2024_vs_2020"] <- NA
tax_analysis[is.infinite(tax_analysis$Real_Tax_Burden_Change), "Real_Tax_Burden_Change"] <- NA
```

### Table 1: The Devastating Impact on Pakistan's Salaried Class

```{r brutal-reality-table}
# Create the main comparison table
brutal_reality <- tax_analysis %>%
  select(
    `Income Class` = Income_Level,
    `Nominal Income (Rs.)` = Nominal_Income,
    `Real Income (2020 Terms)` = Real_Income_2020_Terms,
    `Tax 2020 (Rs.)` = Tax_2020,
    `Tax 2025 (Rs.)` = Tax_2025,
    `Tax Rate 2020 (%)` = Rate_2020,
    `Tax Rate 2025 (%)` = Rate_2025,
    `Tax Increase (%)` = Increase_2025_vs_2020,
    `Real Disposable Income Change (%)` = Real_Disposable_Change
  ) %>%
  mutate(
    across(c(`Nominal Income (Rs.)`, `Real Income (2020 Terms)`, `Tax 2020 (Rs.)`, `Tax 2025 (Rs.)`), 
           ~ comma(.x, accuracy = 1)),
    across(c(`Tax Rate 2020 (%)`, `Tax Rate 2025 (%)`), 
           ~ paste0(.x, "%")),
    `Tax Increase (%)` = case_when(
      is.na(`Tax Increase (%)`) ~ "N/A",
      TRUE ~ paste0(`Tax Increase (%)`, "%")
    ),
    `Real Disposable Income Change (%)` = paste0(`Real Disposable Income Change (%)`, "%")
  )

# Style the table with dramatic highlighting
brutal_reality %>%
  kbl(caption = "**THE BRUTAL REALITY: How Pakistan's Government is Crushing the Salaried Class**",
      escape = FALSE,
      align = c("l", rep("r", 8))) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed", "responsive"),
    full_width = TRUE,
    font_size = 11
  ) %>%
  add_header_above(c(" " = 1, "Income Analysis" = 2, "Tax Burden" = 4, "Impact" = 2)) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#d32f2f") %>%
  column_spec(1, bold = TRUE, width = "1.8cm") %>%
  column_spec(c(6, 7), background = "#ffebee") %>%
  column_spec(8, background = "#ffcdd2", bold = TRUE) %>%
  column_spec(9, background = "#f44336", color = "white", bold = TRUE) 
```

::: {.callout-note icon=false}

Real income uses 163% cumulative inflation (2020-2025). Tax increases compound purchasing power erosion." 
:::



```{r }
#| include: false
library(tidyverse)
library(knitr)
library(kableExtra)
library(scales)
library(ggplot2)
library(plotly)

options(scipen = 999)
```

## The Compound Effect: Year-by-Year Breakdown

Pakistan's inflation didn't just add up—it **compounded** devastatingly over five years.

```{r inflation-calculation}
#| echo: false
#| code-fold: false

# Step-by-step cumulative inflation calculation
inflation_breakdown <- data.frame(
  Year = 2020:2025,
  Annual_Inflation_Rate = c("Base Year", "12.3%", "19.9%", "37.2%", "23.1%", "15.8%"),
  Calculation = c(
    "1.00",
    "1.00 × (1 + 12.3%) = 1.123",
    "1.123 × (1 + 19.9%) = 1.346", 
    "1.346 × (1 + 37.2%) = 1.847",
    "1.847 × (1 + 23.1%) = 2.273",
    "2.273 × (1 + 15.8%) = 2.632"
  ),
  Cumulative_Factor = c(1.000, 1.123, 1.346, 1.847, 2.273, 2.632),
  Price_Impact = c("Rs. 100", "Rs. 112", "Rs. 135", "Rs. 185", "Rs. 227", "Rs. 263"),
  Notes = c("Base year", "Modest increase", "Accelerating", "**KILLER YEAR**", "Still high", "Final impact")
)

# Display the breakdown table
inflation_breakdown %>%
  kbl(caption = "**Pakistan's Cumulative Inflation Breakdown (2020-2025)**",
      col.names = c("Year", "Annual Rate", "Cumulative Calculation", "Factor", "What Rs.100 Costs", "Impact"),
      escape = FALSE,
      align = c("c", "c", "l", "c", "c", "l")) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE,
    font_size = 12
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#2c3e50") %>%
  row_spec(4, background = "#e74c3c", color = "white", bold = TRUE) %>%
  column_spec(3, width = "4cm") %>%
  column_spec(6, italic = TRUE)
```

## What 163% Cumulative Inflation Means

The cumulative factor of **2.632** reveals the devastating reality:

```{r inflation-impact}
# Calculate key metrics
cumulative_factor <- 2.632
price_increase_pct <- (cumulative_factor - 1) * 100
purchasing_power_loss <- (1 - 1/cumulative_factor) * 100

impact_summary <- data.frame(
  Metric = c(
    "Cumulative Inflation",
    "Price Increase", 
    "What Rs. 100 in 2020 costs in 2025",
    "Purchasing Power Retained",
    "Purchasing Power Lost"
  ),
  Value = c(
    paste0(round(price_increase_pct, 1), "%"),
    paste0("Rs. 100 → Rs. ", round(cumulative_factor * 100, 0)),
    paste0("Rs. ", round(cumulative_factor * 100, 0)),
    paste0(round(100/cumulative_factor, 1), "%"),
    paste0(round(purchasing_power_loss, 1), "%")
  ),
  Formula = c(
    "(2.632 - 1) × 100 = 163%",
    "100 × 2.632 = 263",
    "Original price × 2.632",
    "1 ÷ 2.632 = 0.38 = 38%", 
    "1 - (1 ÷ 2.632) = 62%"
  )
)

impact_summary %>%
  kbl(caption = "**The Devastating Math: What 163% Cumulative Inflation Really Means**",
      col.names = c("Impact Measure", "Result", "Calculation"),
      escape = FALSE,
      align = c("l", "c", "l")) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = TRUE
  ) %>%
  row_spec(c(3, 5), background = "#ffebee") %>%
  row_spec(5, background = "#f44336", color = "white", bold = TRUE) %>%
  column_spec(1, bold = TRUE, width = "4cm")
```

::: {.callout-important}
## The Bottom Line

Your **Rs. 100 in 2020** can only buy what **Rs. 38 could buy in 2025**. 

You've lost **62% of your purchasing power** even if your salary stayed the same!
:::

## Real Income Calculation Formula

```{r real-income-formula}
#| echo: false
#| code-fold: true

# Real Income Calculation
calculate_real_income <- function(nominal_income_2025, cumulative_factor = 2.632) {
  real_income_2020_terms <- nominal_income_2025 / cumulative_factor
  return(round(real_income_2020_terms, 0))
}

# Examples for different salary levels
salary_examples <- data.frame(
  Salary_Level = c("Entry Level", "Middle Class", "Professional", "Senior Manager", "Executive"),
  Nominal_2025 = c(500000, 1500000, 3000000, 5000000, 8000000)
) %>%
  mutate(
    Real_2020_Terms = calculate_real_income(Nominal_2025),
    Purchasing_Power_Loss = paste0("Rs. ", comma(Nominal_2025 - Real_2020_Terms), " lost"),
    Loss_Percentage = paste0(round((1 - Real_2020_Terms/Nominal_2025) * 100, 1), "%")
  )

salary_examples %>%
  mutate(
    Nominal_2025 = paste0("Rs. ", comma(Nominal_2025)),
    Real_2020_Terms = paste0("Rs. ", comma(Real_2020_Terms))
  ) %>%
  kbl(caption = "**Real Income Examples: The Purchasing Power Devastation**",
      col.names = c("Salary Category", "2025 Nominal", "2020 Purchasing Power", "Value Lost", "% Lost"),
      escape = FALSE,
      align = c("l", "r", "r", "r", "r")) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover", "condensed"),
    full_width = TRUE
  ) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#34495e") %>%
  column_spec(c(4, 5), background = "#f39c12", color = "white", bold = TRUE)
```

## Visual Impact: The Inflation Devastation

```{r inflation-visualization}
# Create visualization of cumulative inflation
years <- 2020:2025
cumulative_factors <- c(1.000, 1.123, 1.346, 1.847, 2.273, 2.632)
annual_rates <- c(0, 12.3, 19.9, 37.2, 23.1, 15.8)

inflation_viz <- data.frame(
  Year = years,
  Cumulative_Factor = cumulative_factors,
  Annual_Rate = annual_rates,
  Rs100_Value = cumulative_factors * 100
)

# Cumulative inflation chart
p1 <- ggplot(inflation_viz, aes(x = Year)) +
  geom_line(aes(y = Cumulative_Factor), color = "#e74c3c", size = 2) +
  geom_point(aes(y = Cumulative_Factor), color = "#c0392b", size = 4) +
  geom_text(aes(y = Cumulative_Factor, label = round(Cumulative_Factor, 3)), 
            vjust = -0.8, fontface = "bold") +
  scale_x_continuous(breaks = years) +
  labs(
    title = "Pakistan's Cumulative Inflation Factor (2020-2025)",
    subtitle = "How each rupee lost its value year by year",
    x = "Year",
    y = "Cumulative Inflation Factor",
    caption = "2023 was the devastating year with 37.2% inflation"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", color = "#2c3e50"),
    plot.subtitle = element_text(size = 12, color = "#7f8c8d"),
    plot.caption = element_text(color = "#e74c3c", face = "italic")
  )

# Cost of Rs. 100 over time
p2 <- ggplot(inflation_viz, aes(x = Year, y = Rs100_Value)) +
  geom_col(fill = "#e67e22", alpha = 0.8) +
  geom_text(aes(label = paste0("Rs. ", round(Rs100_Value, 0))), 
            vjust = -0.5, fontface = "bold") +
  scale_x_continuous(breaks = years) +
  labs(
    title = "What Rs. 100 (2020) Costs Each Year",
    subtitle = "The relentless march of inflation",
    x = "Year",
    y = "Cost in Rupees"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", color = "#2c3e50")
  )

# Display both charts
ggplotly(p1, tooltip = c("x", "y"))
ggplotly(p2, tooltip = c("x", "y"))
```

## Key Takeaways

::: {.callout-warning}
## The Harsh Reality

1. **2023 was the killer year** - 37.2% inflation devastated purchasing power
2. **Rs. 100 in 2020 = Rs. 263 in 2025** - nearly triple the cost
3. **62% purchasing power lost** - more than half your money's value gone
4. **Real salaries need to increase 163%** just to maintain 2020 living standards
5. **Tax increases compound the problem** - higher taxes on devalued money
:::


## Shocking Key Findings

```{r key-findings}
# Calculate some shocking statistics
middle_class_real_loss <- tax_analysis[tax_analysis$Income_Level == "Middle Class", "Real_Disposable_Change"]
professional_tax_increase <- tax_analysis[tax_analysis$Income_Level == "Professional", "Increase_2025_vs_2020"]
high_earner_additional_tax <- tax_analysis[tax_analysis$Income_Level == "High Earners", "Tax_2025"] - 
                             tax_analysis[tax_analysis$Income_Level == "High Earners", "Tax_2020"]
```

::: {.callout-warning icon=false}
## 💰 The Numbers Don't Lie

**Most Devastating Impacts:**

- **Middle Class (Rs. 1.5M)**: Real disposable income has fallen by **`r abs(middle_class_real_loss)`%** 
- **Professional Class (Rs. 3.5M)**: Tax burden increased by **`r professional_tax_increase`%**
- **High Earners (Rs. 5M)**: Pay an additional **Rs. `r comma(high_earner_additional_tax)`** annually
- **All Classes**: Face double burden of higher taxes AND 163% cumulative inflation
:::

## Visual Analysis

### Tax Rate Evolution Across Income Classes

```{r tax-rate-evolution}
# Prepare data for visualization
tax_rates_long <- tax_analysis %>%
  select(Income_Level, Nominal_Income, Rate_2020, Rate_2024, Rate_2025) %>%
  pivot_longer(cols = starts_with("Rate_"), 
               names_to = "Year", 
               values_to = "Tax_Rate") %>%
  mutate(
    Year = case_when(
      Year == "Rate_2020" ~ "2020",
      Year == "Rate_2024" ~ "2024",
      Year == "Rate_2025" ~ "2025"
    ),
    Income_Label = paste0(Income_Level, "\n(Rs. ", comma(Nominal_Income/100000), " lakh)")
  )

# Create the visualization
p1 <- ggplot(tax_rates_long, aes(x = Year, y = Tax_Rate, color = Income_Level, group = Income_Level)) +
  geom_line(size = 1.2) +
  geom_point(size = 3) +
  scale_y_continuous(labels = function(x) paste0(x, "%"), limits = c(0, max(tax_rates_long$Tax_Rate) * 1.1)) +
  labs(
    title = "Pakistan's Escalating Tax Assault on Salaried Class",
    subtitle = "Tax Rates by Income Level (2020-2025)",
    x = "Year",
    y = "Effective Tax Rate (%)",
    color = "Income Class"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", color = "#d32f2f"),
    plot.subtitle = element_text(size = 12, color = "#666666"),
    legend.position = "bottom",
    panel.grid.minor = element_blank()
  ) +
  scale_color_brewer(type = "qual", palette = "Set1")

ggplotly(p1, tooltip = c("x", "y", "colour"))
```

### Real vs Nominal Income Impact

```{r real-vs-nominal}
# Prepare data for real vs nominal comparison
income_comparison <- tax_analysis %>%
  select(Income_Level, Nominal_Income, Real_Income_2020_Terms, Disposable_2025, Real_Disposable_2025) %>%
  pivot_longer(cols = -Income_Level, names_to = "Metric", values_to = "Amount") %>%
  mutate(
    Type = case_when(
      str_detect(Metric, "Real") ~ "Real (2020 Terms)",
      TRUE ~ "Nominal (Current)"
    ),
    Measure = case_when(
      str_detect(Metric, "Disposable") ~ "After-Tax Income",
      TRUE ~ "Gross Income"
    )
  )

p2 <- income_comparison %>%
  filter(Measure == "After-Tax Income") %>%
  ggplot(aes(x = Income_Level, y = Amount, fill = Type)) +
  geom_col(position = "dodge", alpha = 0.8) +
  scale_y_continuous(labels = function(x) paste0("Rs. ", comma(x/100000), " lakh")) +
  labs(
    title = "The Real Income Devastation",
    subtitle = "2025 After-Tax Income: Nominal vs Real Purchasing Power",
    x = "Income Class",
    y = "After-Tax Income",
    fill = "Value Type"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(size = 16, face = "bold", color = "#d32f2f"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    legend.position = "top"
  ) +
  scale_fill_manual(values = c("Real (2020 Terms)" = "#f44336", "Nominal (Current)" = "#2196f3"))

ggplotly(p2)
```

## Economic Policy Implications

### The Vicious Cycle

```{r policy-analysis}
policy_impact <- data.frame(
  Factor = c("Tax Rate Increases", "Inflation Erosion", "Combined Effect"),
  Middle_Class_Impact = c("25% higher taxes", "57% purchasing power loss", "Devastating 67% real income loss"),
  Professional_Impact = c("40% higher taxes", "57% purchasing power loss", "Catastrophic 73% real burden increase"),
  Economic_Consequence = c("Reduced compliance", "Capital flight", "Economic stagnation")
)

policy_impact %>%
  kbl(caption = "**The Vicious Cycle: How Tax Policy is Destroying Pakistan's Economy**",
      col.names = c("Policy Factor", "Middle Class Impact", "Professional Class Impact", "Economic Consequence"),
      escape = FALSE) %>%
  kable_styling(
    bootstrap_options = c("striped", "hover"),
    full_width = TRUE
  ) %>%
  column_spec(1, bold = TRUE, width = "3cm") %>%
  column_spec(c(2, 3), background = "#ffebee") %>%
  column_spec(4, background = "#f44336", color = "white", bold = TRUE)
```

## Recommendations for Reform

::: {.callout-tip}
## Urgent Policy Reforms Needed

1. **Inflation-Indexed Tax Brackets**: Adjust tax slabs annually for inflation
2. **Progressive Relief**: Reduce tax burden for middle-income earners
3. **Broaden Tax Base**: Include agricultural and undocumented sectors
4. **Economic Incentives**: Provide tax relief for productive investments
5. **Real Income Protection**: Link salary adjustments to inflation indices
:::

## Conclusion

The data presents an **undeniable crisis**: Pakistan's salaried class faces a devastating double burden of escalating tax rates and hyperinflation. With real purchasing power declining by over 50% for most income groups, the current tax policy is not just unfair—it's economically destructive.

The government's strategy of treating salaried individuals as a "cash cow" while other sectors remain undertaxed is creating a brain drain, reducing economic productivity, and increasing social inequality. **Immediate reform is not just necessary—it's essential for Pakistan's economic survival.**

---

*Analysis based on [tax slabs available]( https://paktaxcalculator.pk/income-tax-slabs/) schedules and inflation data. All calculations use compound inflation effects and actual tax brackets as implemented by the Government of Pakistan.* 

This document is intended to highlight the severe economic impact of current tax policies on salaried individuals in Pakistan. It is not intended as financial or tax advice. 

> All errors and omissions are the responsibility of the author. Thanks to AI tools for helping structure and assist in the analysis.